---
title: "EditR v`r editrversion` report"
output: 
  html_document: 
    toc: yes
params:
  null.m.params: NA
  guide: NA
  guide.coord: NA
  editing.df: NA
  p.val.cutoff: NA
  input.seq: NA
  sangs: NA
  sangs.filt: NA
  base.info: NA
  editrversion: NA
---

```{r echo = FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```


```{r, echo = FALSE}
null.m.params <- params$null.m.params
guide <- params$guide
guide.coord <- params$guide.coord
editing.df <- params$editing.df
p.val.cutoff <- params$p.val.cutoff
input.seq <- params$input.seq
sangs <- params$sangs
sangs.filt <- params$sangs.filt
base.info <- params$base.info
editrversion <- params$editrversion
```



## Data QA

### Filtering data

What the data looked like prefiltering:

```{r}
ggplot(sangs, aes(x = index, y = Tot.area)) +
       geom_rect(xmin = guide.coord$start, xmax = guide.coord$end, ymin = 0, ymax = Inf, fill = "lightgrey") +
        geom_line() +
        labs(x = "Base position",
             y = "Total peak area at base position (A + T + C + G)",
             title = "Unfiltered data, total peak area")
```


and the post filtering signal / noise plot:

```{r}
## doing some data rearrangement for plotting
      
catch <- sangs.filt %>% dplyr::select(A.area:T.area, A.perc:T.perc, base.call, index)
catch %<>% gather( key = base, value = value,
                   A.area:T.area, A.perc:T.perc) %>%
  separate(col = base, into = c("base", "measure"))

noise <- catch %>% 
  group_by(index, measure) %>% 
  filter(base != base.call) %>%
  summarize(noise = sum(value))
signal <- catch %>%
  group_by(index, measure) %>%
  filter(base == base.call) %>% 
  summarize(signal = sum(value))

summary.df <- left_join(noise,signal) %>%
  gather(key = type, value = value, noise, signal)

sangs.plot <- sangs.filt %>% dplyr::select(index, base.call) 
sangs.plot %<>% left_join(summary.df)

## now acutally making the plot

sangs.plot$type <- ordered(sangs.plot$type, levels = c("signal", "noise"))

max.height <- filter(sangs.plot, measure == "area") %>% select(value) %>% max

guide.region <- data.frame(index = c(guide.coord$start, guide.coord$end,
                                     guide.coord$end, guide.coord$start), 
                           value = c(0, 0, max.height, max.height))

p <- sangs.plot %>% filter(measure == "area") %>%
  ggplot(aes(x = index, y = value)) +
  geom_polygon(data = guide.region, fill = "lightgrey") +
  geom_line(aes(color = type)) +
  scale_color_manual(values = c("#5e3c99", "#e66101")) +  
  labs(title = "Filtered data signal and noise total area",
       x = "Position",
       y = "Peak Area") + 
  theme(legend.title = element_blank())
ggplotly(p)
```


### Percent noise peak area

```{r}
## doing some data rearrangement for plotting
      
      catch <- sangs.filt %>% dplyr::select(A.area:T.area, A.perc:T.perc, base.call, index)
      catch %<>% gather( key = base, value = value,
                         A.area:T.area, A.perc:T.perc) %>%
        separate(col = base, into = c("base", "measure"))
      
      # splitting the catch dataframe into either signal and noise, and calculating the 
      # total noise area or total noise percent
      noise <- catch %>% 
        group_by(index, measure) %>% 
        filter(base != base.call) %>%
        summarize(noise = sum(value))
      signal <- catch %>%
        group_by(index, measure) %>%
        filter(base == base.call) %>% 
        summarize(signal = sum(value))
      
      signal.noise.df <- left_join(noise,signal) %>%
        gather(key = type, value = value, noise, signal)
      
      # making the plotting df
      sangs.plot <- sangs.filt %>% dplyr::select(index, base.call) 
      sangs.plot %<>% left_join(signal.noise.df)
      sangs.plot$type <- ordered(sangs.plot$type, levels = c("signal", "noise"))
      
      sangs.plot %>% filter(measure == "perc", type == "noise") %>%
        ggplot(aes(x = index, y = value)) +
        geom_area(aes(fill = type)) +
        scale_fill_manual(values = c("#e66101")) + 
        annotate("rect", xmin=guide.coord$start, xmax=guide.coord$end,
                 ymin=0, ymax=Inf, alpha=1/5, fill="black") +
        guides(fill = FALSE) + 
        labs(title = "Percent peak area noise",
             x = "Position",
             y = "Percent peak area noise")
```


### Base information

Here's information about the signal of each base, the critical percent value where any higher value would be called as significant, and Filliben's correlation for how well the noise was modelled by the zero adjusted gamma distribution. 

```{r}
temp <- base.info
names(temp) <- c("Base", "Average percent signal", 
                 "Average peak area",  "Critical percent value",
                 "model mu",  "Fillibens correlation")
row.names(temp) <- NULL

knitr::kable(temp)
```


## Predicted editing 


### Editing bar plot

```{r}
      edit.long <- editing.df %>% gather(key = focal.base, value = value, 
                                         A.area:T.area, A.perc:T.perc, T.pval:A.pval) %>%
        separate(col = focal.base, into = c("focal.base", "measure"))
      
      p.a <- makeEditingBarPlot(edit.long = edit.long, null.m.params = null.m.params$a,
                                base = "A", pval = p.val.cutoff, editing.df)
      p.c <- makeEditingBarPlot(edit.long, null.m.params$c,
                                base = "C", pval = p.val.cutoff, editing.df)
      p.g <- makeEditingBarPlot(edit.long, null.m.params$g,
                                base = "G", pval = p.val.cutoff, editing.df)
      p.t <- makeEditingBarPlot(edit.long, null.m.params$t,
                                base = "T", pval = p.val.cutoff, editing.df)
      
      grid.arrange(p.a, p.c, p.g, p.t)
```

### Editing table plot

```{r}
edit.long <- editing.df %>% gather(key = focal.base, value = value, 
                                         A.area:T.area, A.perc:T.perc, T.pval:A.pval) %>%
        separate(col = focal.base, into = c("focal.base", "measure"))
      
edit.spread <- edit.long %>% spread(key = measure, value = value)

edit.color <- edit.spread %>% filter(pval < p.val.cutoff)

if(any(edit.color$adj.perc != 100)){
edit.spread %>%
        ggplot(aes(x = as.factor(index), y = focal.base)) + 
        geom_tile(data = edit.color, aes(fill = adj.perc)) + 
        geom_text(aes(label = round(perc, 0)), angle = 0, size = 5) +   
        guides(fill = FALSE) + 
        scale_fill_continuous(low = "#f7a8a8", high = "#9acdee") + 
        scale_x_discrete(position = "top", labels = editing.df$guide.seq) + 
        labs(x = NULL, y = NULL) + 
        theme(axis.ticks = element_blank(),
              axis.text=element_text(size=16),
              plot.title = element_text(hjust = 0, size = 16),
              plot.margin=unit(c(0,0,0,2), "cm"), #c(top, bottom, left, right)
              panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
              plot.background = element_rect(fill = "transparent",colour = NA)
        ) +
        coord_fixed(1)} else
      {edit.spread %>%
        ggplot(aes(x = as.factor(index), y = focal.base)) + 
        geom_tile(data = edit.color, fill = "#9acdee") + 
        geom_text(aes(label = round(perc, 0)), angle = 0, size = 5) +   
        guides(fill = FALSE) + 
        scale_x_discrete(position = "top", labels = editing.df$guide.seq) + 
        labs(x = NULL, y = NULL) + 
        theme(axis.ticks = element_blank(),
              axis.text=element_text(size=16),
              plot.title = element_text(hjust = 0, size = 16),
              plot.margin=unit(c(0,0,0,2), "cm"), #c(top, bottom, left, right)
              panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
              plot.background = element_rect(fill = "transparent",colour = NA)
        ) +
        coord_fixed(1)}
```


### Table of editing results

<!-- Here are just the positions that showed editing: -->

<!-- ```{r} -->
<!-- options(scipen=999) -->
<!-- edit.spread %>% filter(pval < p.val.cutoff, focal.base != guide.seq) %>% -->
<!--   select(index, guide.position, guide.seq, base.call, focal.base, area, perc, pval) %>%  -->
<!--   knitr::kable(digits = c(0,0,0,0,0,2,999), -->
<!--                col.names = c("Sanger position", "Guide position","Guide sequence", "Sanger base call", -->
<!--                              "Focal base", "Focal base peak area", -->
<!--                              "Focal base percent area", "p value")) -->
<!-- ``` -->

Here's the entire guide region

```{r}
edit.spread %>% arrange(guide.position) %>% select(index,guide.position, guide.seq, base.call, focal.base, perc, pval) %>% 
  mutate(perc = format(perc, digits = 2), signficant = ifelse(pval < p.val.cutoff, yes = "*", no = "")) %>%
    knitr::kable(digits = c(0,0,0,0,0,2,999, 0),
               col.names = c("Sanger position", "Guide position","Guide sequence", "Sanger base call",
                             "Focal base", "Focal base peak area", "p value", ""))
```





## For use in R

If you want to work with the results in R, here is output that you can copy and paste in your terminal to get:

The base information: 

```{r, echo=FALSE, comment = NA}
temp <- base.info
row.names(temp) <- NULL
dput(temp)
```

the data.frame that contains information on the guide region:

```{r echo = FALSE, comment = NA}
dput(editing.df)
```



*Report generated using EditR v`r editrversion`*

